\TheTitle

### 2015-08-17 linvis.tex
Editing: Introduction and Related Work

Maintainers of older version of the Linux kernel must sift through thousands of
commits with tools that are unable to filter and effectively visualize projects
that are of the scale of the kernel. Tools like Gitk provide users with the
full view of the DAG, showing all merges and commits. For smaller projects,
this information is useful for showing the relation between various branches.
The DAG becomes a tangled mess (\ref{fig:gitk}) of commits, merges, and the
links between them with large modular projects, like the Linux kernel. This
makes it difficult to derive an explanation of the changes made. In other
cases, the kernel is simply too large for the system to generate a
visualization for. Github provides a DAG view for most projects, but is unable
to display the visualization for projects as large as the Linux kernel.

@@

Git log, Gitk, and other similar tools only store the authored date of a
commit. A commit may be merged into the project at any time, regardless of when
the commit was authored, making date-range queries meaningless. Performing a
date-range query using other tools over the span of a release will contain
results that are for previous versions of the kernel, and may not contain all
of the results for that kernel if patches were merged after the release date.

German et al. recorded the metadata of every commit and merge in the Linux git
repository starting in 2011 \cite{German}, including chronological data that is
not stored by git. We are able to use this information to generate sets of
trees instead of a DAG. The root of each tree is the top-level merge performed
by Linus, which merges branches or sets of commits into the kernel.

In this paper, we present the design decisions behind a web-based
tool\footnote{\url{http://li.turingmachine.org}} built around a tree-based
model instead of the DAG.  We demonstrate the advantages of this model over the
DAG two main use-cases.  Our tool provides information about the location of a
given commit in the respective merge tree, the files edited, the modules
edited, and the commit message in a given commit or merge.  The tool allows
users to apply various filters, including the release, a keyword or phrase from
the log preview, the name of the author, or the commit id. The user can request
all the top-level merges containing a commit or merge that matches the query,
or all commits and merges that match the query.

Our core model provides advantages to the DAG model used by other tools.
Breaking the commits and merges by merge tree removes the commits and merges of
other components of the kernel from the visualization, providing a clearer
picture of what is happening in a given merge. This model provides additional
support to edited file information. Furthermore, we are able to understand the
changes being made in the order they are brought into the kernel.

All top-level merges to a given version of the kernel are merged between the
start and release date. Our tree construction allows us to use the top-level
merges to find all merges and commits that have an effect on a given version of
the kernel. The model allows us to better provide mechanisms for performing
date-range queries, providing access to all commits that have an effect on a
given release, even if the commit comes before the start date, or after the
release date.

\section{Related Work}

To our knowledge, no repository visualization systems using a tree-based model
are being used. This may be due to the fact that more information than what
is stored by the normal tools is required to create the trees, so servers must
sit and watch as the repository changes over time. Many of the visualization
tools are designed for a general-purpose approach, where any repository is
directly usable, without additional information. Polling for this additional
information makes the application specific to a single repository, or project.

\subsection{Gitk}
Gitk, Gitg, and other similar git repository visualizing tools use the DAG as
the central organization. For smaller projects, this model is sufficient for
providing meaningful information, but in larger projects like the Linux kernel,
the branches become too tangled and too deep to derive any meaning from the
DAG. The DAG model allows users to go to the next commit in the chain, go to
the previous commit in the chain, or jump to the parent merge. This suggest
that Gitk and similar tools are designed for users working with a bottom-to-top
approach. With this model, it is difficult for users to see what other commits
are involved with a given merge. Gitk only provides file information on
commits, as it is unable to quickly find all files edited by a merge. It is up
to the user to remember what files were edited in each of the commits involved
in a given merge. In some of the larger merges, some containing over 1500
edited files, this task become nearly impossible.
% cid: 73287a43cc79ca06629a88d1a199cd283f42456a, 1800 commits, 1500 files
Other DAG visualization systems, like the GitHub viewer (\ref{fig:gitfail}),
are unable to provide a visualization for projects that are as large as the
kernel.

@@

\subsection{Treemap visualizer}
University of Maryland presented a treemap visualization
tool\footnote{\url{https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html}},
pictured in figure \ref{fig:treemap}, for
displaying all the file structure of a given version. Treemaps are good for
displaying the topology of shallow, wide trees, making it a good candidate for
displaying file systems. Treemaps are limited in what information can be
displayed. The information within the treemap does not provide clear insight on
what changes were made to the files in the given release. The treemap
visualization only provides information on where the file is located within the
filesystem. File systems and the merge tree of the kernel share many
similarities, in both cases, the representative trees are wide and shallow,
suggesting that a treemap could potentially work to visualize the commit and
merge structure of the kernel.

The treemap is unable to take into account multiple merges at the same level
with the same name. There is an additional dimension, in that the commits and
merges are ordered in time. The treemap design is unable to show the ordering
of the merges and commits.

@@

Finally, the treemap does not provide any additional information about the
contents of each cell, only that the cell exists. The user is unable to see
their current position within the map, or if we assume that their current
position is the larges surrounding box, they can no-longer see how their
current position interacts with the rest of the system.
%% https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html
%% Find paper for this maybe?

@@

\section{Design}

The goal is to build a tool that make navigation of the kernel commit
information simpler, while providing an explanation of the changes occurring at
each commit and merge. Part of making the information usable is to narrow the
results down into only the information that is relevant to what the user is
looking for. We designed the tool with two use-cases in mind, though a user may
switch roles many times as they work.

@@

\subsection{Command-line Shell}

Our initial intuition was to work with the commits and merges as trees, like in
a file structure. There are many similarities in the structure of a file system
and the structure of the commit information within the kernel. The primary
similarity is that they are both trees, and both trees are generally fairly
shallow and very wide. File systems have been around for quite some time, and
there are many tools for visualizing and working withe the file structure. The
first tool was the command-line. We built a small command-line shell
(\ref{fig:shell}) that allowed navigation through the commit and merges instead
of files and directory. The shell design permitted a user to move through the
merges like they would through directories. They could use the cat command to
see the preview of the log, or the more command to see the preview and what
files were edited in the commit and how many lines were added and removed from
each file.

@@

\subsection{Web-based Tool}

There are multiple reasons for building the system into a web-based tool.
Putting the tool on the web enables users to use the system without having to
install any software, making the system more accessible and platform
independent.A web-based tool can utilize more interactive means of navigation,
and provide a better explanation of what each commit or merge may contain than
the command-line variant.

@@

\section{Implementation Details}

We built a we-based tool to provide information about the kernel.
We chose to use Nginx as the main web server, postgresql as the dbms, and Flask
as the framework for generating the dynamic content.

>>>>

Maintainers of older version of the Linux kernel must sift through thousands of
commits with tools that are unable to filter and effectively visualize projects
that are the scale of the kernel. Tools like Gitk provide users with the
full view of the DAG, showing all of the merges and commits. For smaller
projects, this information is useful, showing the relation between various
branches. The DAG becomes a tangled mess (\ref{fig:gitk}) of commits, merges,
and the links between them with large modular projects, like the Linux kernel,
making it difficult to derive an explanation of the changes made. In other
cases, the kernel is simply too large for the system to generate a
visualization for. Github provides a DAG view for most projects, but is unable
to display the visualization for projects as large as the Linux kernel.

@@

Git log, Gitk, and other similar tools only store the authored date of a
commit. A commit may be merged into the project at any time, regardless of when
the commit was authored, making date-range queries meaningless. Performing a
date-range query using these tools over the span of a release will contain
results that are for previous versions of the kernel or may not contain all
of the results for that kernel if patches were merged after the release date.

German et al. recorded the metadata of every commit and merge in the Linux git
repository starting in 2011 \cite{German}, including chronological data that is
not stored by git. We are able to use this information to generate sets of
trees instead of a DAG. Each tree is rooted by the top-level merge, performed
by Linus, which merges the branches or a set of commits into the kernel.

In this paper, we present our design decisions behind a web-based
tool\footnote{\url{http://li.turingmachine.org}} built around a tree-based
model instead of a DAG. Our tool provides information about the location of a
given commit in the respective merge tree, the files edited, the modules
edited, and the commit message for a given commit or merge. The tool allows
users to apply various filters, including the release version, along with a
keyword or phrase from the log preview, the name of the author, or the commit
id. The user can request all the top-level merges containing a commit or merge
that matches the query, or all commits and merges that match the query.

Our core model provides advantages to the DAG model used by other tools.
Breaking the commits and merges by merge tree removes commits and merges of
other components of the kernel from the visualization, providing a clearer
picture of what is happening in a given merge and providing additional
support for edited file information. Furthermore, we are able to understand the
changes being made, in the order they are brought into the kernel.

All top-level merges to a given version of the kernel are merged between the
start and release date. Our tree construction allows us to use the top-level
merges to find all merges and commits that have an effect on a given version of
the kernel. The model allows us to better provide mechanisms for performing
date-range queries, providing access to all commits that have an effect on a
given release, even if the commit comes before the start date or after the
release date.

\section{Related Work}
To our knowledge, we know of no repository visualization systems using a
tree-based model. This may be due to the fact that more information is required
to generate the trees than what is stored by git. A server must sit and monitor
the repository as it changes over time. Many of the visualization tools are
designed for a general-purpose approach, where any repository is directly
usable without additional information. Polling for this additional information
makes the application specific to a single repository or project.

\subsection{Gitk}
Gitk, Gitg, and other similar git repository visualizing tools use the DAG as
the central organization model. For smaller projects, this model is sufficient
for providing meaningful information, but in larger projects like the Linux
kernel, the branches become too tangled and too deep to derive any meaning from
the DAG. The DAG model allows users to go to the next commit in the chain, go
to the previous commit in the chain, or jump to the parent merge. This suggest
that Gitk and similar tools are designed for users working with a bottom-to-top
approach. With this model, it is difficult for users to see what other commits
are involved with a given merge. Gitk only provides file information on
commits, as it is unable to quickly find all files edited by the commits within
a merge. It is up to the user to remember what files were edited in each of the
commits involved in a given merge. In some of the larger merges, some
containing over 1500 edited files, this task become nearly impossible.
% cid: 73287a43cc79ca06629a88d1a199cd283f42456a, 1800 commits, 1500 files
Other DAG visualization systems, like the GitHub viewer (\ref{fig:gitfail}),
are unable to provide a visualization for projects that are as large as the
kernel.

@@

\subsection{Treemap visualizer}
University of Maryland presented a treemap visualization
tool\footnote{\url{https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html}},
pictured in figure \ref{fig:treemap}, for
displaying the file structure of a given version. Treemaps are good for
visualizing the topology of shallow, wide trees, making it a good candidate for
showing file systems. Treemaps are limited in what information can be
displayed. The information within the treemap does not provide clear insight on
what changes were made to the files in the given release, only providing
information on where the file is located within the file system.
File systems and the merge tree of the kernel share many similarities, in both
cases, they represent trees that are wide and shallow, suggesting that a
treemap could potentially work to visualize the commit and merge structure of
the kernel.

The treemap is unable to take into account multiple commits behind the same
merge with the same name. Commits and merges have an ordering in which they
were added to the kernel, which cannot be represented by a treemap, but is
important and can be used to assist in building the tree.

@@

Finally, the treemap does not provide any additional information about the
contents of each cell, only that the cell exists and where it exists in the
tree. The user is unable to see their current position within the map, or if we
assume that their current position is the largest surrounding box, they can
no longer see how their current position interacts with the rest of the
system.
%% https://www.cs.umd.edu/hcil/millionvis/Treemap_Visualization_of_the_Linux_Kernel_2_5_33.html
%% Find paper for this maybe?

@@

\section{Design}
The goal is to build a tool that makes navigation of the kernel commit
information simpler, while providing an explanation of the changes occurring at
each commit and merge. Part of making the information usable is to narrow the
results down into only the information that is relevant to what the user is
looking for. We designed the tool with two use-cases in mind, though a user may
switch roles many times as they work.

@@

\subsection{Command-line Shell}
Our initial intuition was to work with the commits and merges as trees, like in
a file structure. There are many similarities in the structure of a file system
and the structure of the commit information within the kernel. The primary
similarity is that they are both trees, and both trees are generally fairly
shallow and very wide. File systems have been around for quite some time, and
there are many tools for visualizing and working withe the file structure. The
first tool was the command-line. We built a small command-line shell
(\ref{fig:shell}) that allowed navigation through the commit and merges instead
of files and directory. The shell design permitted a user to move through the
merges like they would through directories. They could use the cat command to
see the preview of the log, or the more command to see the preview and what
files were edited in the commit and how many lines were added and removed from
each file.

@@

\subsection{Web-based Tool}
There are multiple reasons for building the system into a web-based tool.
Putting the tool on the web enables users to use the system without having to
install any software, making the system more accessible and platform
independent.A web-based tool can utilize more interactive means of navigation,
and provide a better explanation of what each commit or merge may contain than
the command-line variant.

@@

\section{Implementation Details}
We built a we-based tool to provide information about the kernel.
We chose to use Nginx as the main web server, postgresql as the dbms, and Flask
as the framework for generating the dynamic content.
